---
title: "Bibliometric Analysis of Small Lakes Research (1990–2024)"
author: "Kate Regine Cunanan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

```{r load_data, message=FALSE, warning=FALSE}
# Load libraries
library(bibliometrix)
library(tidyverse)
library(ggplot2)
library(maps)

# Import raw Scopus data
scopus_raw <- convert2df(
  file = "_small lake__.csv [no keyword limitation]",
  dbsource = "scopus",
  format = "csv"
)

# Run bibliometric analysis
bib_analysis <- biblioAnalysis(scopus_raw, sep = ";")
bib_summary <- summary(bib_analysis, k = 10, pause = FALSE)
p1 <- plot(bib_analysis, k = 10, pause = FALSE)
```

## 1. Introduction

Bibliometric analysis is a quantitative method used to evaluate and map the scientific literature of a specific field. This study uses bibliometric analysis to examine the research landscape on **small lakes** from 1990 to 2024, using data retrieved from the Scopus database.

The analysis aims to identify publication trends, leading authors, influential journals, productive countries, and dominant research themes in the field of small lakes research.

## 2. Data and Methods

Data was retrieved from the **Scopus database** using "small lake*" and related terms as search keywords. The raw dataset consisted of 13,181 records spanning from 1898 to 2025.

The dataset was cleaned using the following steps:

- Filtered to include only publications from **1990 to 2024**
- Retained only **Articles and Reviews** as document types
- Removed **72 duplicate records** based on title

The final cleaned dataset comprises **11,391 documents** published across **1,729 sources** by **28,700 authors**.

All analyses and visualizations were performed in R using the `bibliometrix`, `tidyverse`, and `ggplot2` packages.

## 3. Data Cleaning

The raw dataset retrieved from Scopus consisted of **13,181 records** spanning 
from 1898 to 2025. Before proceeding with the analysis, the data was inspected 
and cleaned through the following steps.

### 3.1 Checking for Missing Values

The first step was to check whether key bibliometric fields contained any missing 
values.

```{r missing_values}
colSums(is.na(scopus_raw[, c("AU", "TI", "PY", "SO", "DE", "ID", "AB", "TC", "CR", "C1")]))
```

All key fields returned zero missing values, indicating that the Scopus export 
was complete and no imputation was necessary.

### 3.2 Checking Publication Years

Next, the distribution of publication years was examined to identify any 
anomalous or sparse records.

```{r year_distribution}
table(scopus_raw$PY)
```

The data contained records as far back as 1898, with publication counts remaining 
very sparse prior to 1990. To ensure a meaningful and consistent trend analysis, 
the dataset was filtered to include only publications from **1990 onwards**. 
Additionally, the single record from 2025 was excluded as the year is incomplete 
and would skew the publication trend downward.

### 3.3 Checking for Duplicates

```{r duplicates}
sum(duplicated(scopus_raw$TI))
```

A total of **72 duplicate records** were identified based on document title and 
were removed to avoid inflating author and citation counts.

### 3.4 Checking Document Types

```{r document_types}
table(scopus_raw$DT)
```

The dataset contained 11 document types, including articles, reviews, conference 
papers, book chapters, books, editorials, and others. For bibliometric analysis, 
only **peer-reviewed articles and reviews** were retained, as these are the most 
comparable and citable document types in the literature.

### 3.5 Applying the Cleaning Steps

Based on the inspection above, the following cleaning steps were applied:

- Filter to publications from **1990 to 2024**
- Retain only **Articles** and **Reviews**
- Remove **72 duplicate records** based on title

```{r data_cleaning}
# Filter by year
scopus_clean <- scopus_raw[scopus_raw$PY >= 1990 & scopus_raw$PY <= 2024, ]

# Keep only Articles and Reviews
scopus_clean <- scopus_clean[scopus_clean$DT %in% c("ARTICLE", "REVIEW"), ]

# Remove duplicates
scopus_clean <- scopus_clean[!duplicated(scopus_clean$TI), ]

# Confirm final dataset dimensions
dim(scopus_clean)
```

The cleaned dataset comprises **11,391 documents** ready for bibliometric 
analysis.

## 4. Descriptive Overview

The following key metrics summarize the dataset:

| Metric | Value |
|---|---|
| Timespan | 1990–2024 |
| Total Documents | 11,391 |
| Total Authors | 28,700 |
| Total Sources | 1,729 |
| Annual Growth Rate | 3.53% |
| Average Citations per Document | 33.55 |
| Co-Authors per Document | 4.53 |
| International Co-authorships | 27.71% |

## 5. Visualizations

### 5.1 Annual Scientific Production

```{r annual_production}
p1$AnnualScientProd +
  labs(
    title = "Annual Scientific Production on Small Lakes (1990–2024)",
    subtitle = "Based on Scopus database | n = 11,391 documents",
    x = "Year",
    y = "Number of Articles",
    caption = "Source: Scopus | Bibliometric Analysis"
  )
```

### 5.2 Top 10 Most Productive Authors

```{r top_authors}
p1$MostProdAuthors +
  labs(
    title = "Top 10 Most Productive Authors in Small Lakes Research",
    subtitle = "Based on Scopus database | n = 11,391 documents",
    x = "Number of Articles",
    y = "Authors",
    caption = "Source: Scopus | Bibliometric Analysis"
  )
```

### 5.3 Top 10 Most Relevant Journals

```{r top_journals}
journal_df <- as.data.frame(table(scopus_clean$SO))
colnames(journal_df) <- c("Journal", "Articles")
journal_df <- journal_df[order(-journal_df$Articles), ][1:10, ]

ggplot(journal_df, aes(x = reorder(Journal, Articles), y = Articles)) +
  geom_bar(stat = "identity", fill = "#2171b5") +
  coord_flip() +
  labs(
    title = "Top 10 Most Relevant Journals in Small Lakes Research",
    subtitle = "Based on Scopus database | n = 11,391 documents",
    x = "Journal",
    y = "Number of Articles",
    caption = "Source: Scopus | Bibliometric Analysis"
  )
```

### 5.4 Country Scientific Production

```{r country_map}
library(maps)

country_df <- as.data.frame(bib_analysis$Countries)
colnames(country_df) <- c("Country", "Articles")
country_df$Country <- stringr::str_to_title(country_df$Country)

world_map <- map_data("world")
country_df$Country[country_df$Country == "Usa"] <- "USA"
country_df$Country[country_df$Country == "United Kingdom"] <- "UK"

map_merged <- left_join(world_map, country_df, by = c("region" = "Country"))

ggplot(map_merged, aes(x = long, y = lat, group = group, fill = Articles)) +
  geom_polygon(color = "white", linewidth = 0.1) +
  scale_fill_gradient(
    low = "#c6dbef", high = "#08306b",
    na.value = "gray90",
    name = "Articles"
  ) +
  labs(
    title = "Country Scientific Production on Small Lakes Research",
    subtitle = "Based on Scopus database | n = 11,391 documents",
    caption = "Source: Scopus | Bibliometric Analysis"
  )
```

### 5.5 Top 20 Author Keywords

```{r keywords}
keywords <- scopus_clean %>%
  select(DE) %>%
  filter(!is.na(DE)) %>%
  mutate(DE = tolower(DE)) %>%
  separate_rows(DE, sep = ";") %>%
  mutate(DE = str_trim(DE)) %>%
  filter(DE != "")

top_keywords <- keywords %>%
  count(DE, sort = TRUE) %>%
  slice_head(n = 20)

ggplot(top_keywords, aes(x = reorder(DE, n), y = n)) +
  geom_bar(stat = "identity", fill = "#2171b5") +
  geom_text(aes(label = n), hjust = -0.2, size = 3, color = "gray30") +
  coord_flip() +
  labs(
    title = "Top 20 Author Keywords in Small Lakes Research",
    subtitle = "Based on Scopus database | n = 11,391 documents",
    x = "Keyword",
    y = "Frequency",
    caption = "Source: Scopus | Bibliometric Analysis"
  )
```

### 5.6 Thematic Map

```{r thematic_map, fig.width=12, fig.height=8}
thematic_map <- thematicMap(
  scopus_clean,
  field = "DE",
  n = 250,
  minfreq = 5,
  stemming = FALSE,
  size = 0.5,
  n.labels = 1,
  repel = TRUE
)

plot(thematic_map$map) +
  labs(
    title = "Thematic Map of Small Lakes Research",
    subtitle = "Based on Scopus database | n = 11,391 documents",
    caption = "Source: Scopus | Bibliometric Analysis"
  )
```

### 5.7 Top 10 Most Cited Papers

```{r citation_analysis}
cited_df <- scopus_clean %>%
  select(TI, AU, PY, SO, TC) %>%
  arrange(desc(TC)) %>%
  slice_head(n = 10) %>%
  mutate(
    Label = paste0(word(AU, 1), " (", PY, ")"),
    TI = str_trunc(TI, 50)
  )

ggplot(cited_df, aes(x = reorder(Label, TC), y = TC)) +
  geom_bar(stat = "identity", fill = "#2171b5") +
  geom_text(aes(label = TC), hjust = -0.2, size = 3, color = "gray30") +
  coord_flip() +
  labs(
    title = "Top 10 Most Cited Papers in Small Lakes Research",
    subtitle = "Based on Scopus database | n = 11,391 documents",
    x = "Paper",
    y = "Total Citations",
    caption = "Source: Scopus | Bibliometric Analysis"
  )
```

## 6. Conclusion

This bibliometric analysis examined the scientific landscape of small lake research 
using 11,391 documents retrieved from the Scopus database, covering the period 
from 1990 to 2024. The analysis reveals a field that has grown consistently over 
three decades, with an annual growth rate of 3.53% and a peak publication output 
of 685 documents in 2022 — reflecting a growing recognition of the ecological 
importance of small lakes in the global scientific community.

The United States, China, and Canada emerged as the most productive countries, 
while *Hydrobiologia* and the *Journal of Paleolimnology* were identified as the 
leading publication outlets. The high international co-authorship rate of 27.71% 
and an average of 4.53 co-authors per document suggest that small lake research 
is increasingly collaborative and globally engaged.

Thematic analysis of author keywords revealed that the dominant research themes 
revolve around **climate change**, **eutrophication**, **phytoplankton**, and 
**paleolimnology** — all of which are rooted in the natural sciences. This pattern 
is consistent with the broader observation that small lake research has historically 
been concentrated in ecological and hydrological dimensions, with comparatively 
little attention given to the social and economic aspects of these water bodies.

This gap represents a significant opportunity for future research. Small lakes, 
despite being numerically dominant and biologically active, remain underexplored 
in terms of their socio-economic contributions to surrounding communities — 
including their roles in supporting local livelihoods, fisheries, aquaculture, and 
rural economies. Future studies that bridge the natural and social sciences in the 
context of small lake research would contribute meaningfully to a more holistic 
understanding of these ecosystems and their value to human communities.